### 2024-11-05 Paper.qmd
Edited tantrum sections in paper (background and discussion).
Made changes to accomodate the name differences in the new causes dataframe.

---
title: "The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study"
author:
  - name: Michael R. Gaffney
    orcid: 0000-0002-7552-6942
    corresponding: false
    affiliation:
      - name: Washington State University
        department: Anthropology
  - name: Edward H. Hagen
    orcid: 0000-0002-5758-9093
    email: edhagen@wsu.edu
    corresponding: true
    affiliation:
      - name: Washington State University
        department: Anthropology
execute:
  echo: false
  warning: false
format:
  html:
    toc: true
    toc-expand: 2
    number-sections: true
lightbox: true
editor: visual
tbl-cap-location: bottom
bibliography: signaling.bib
crossref:
  custom:
    - kind: float
      reference-prefix: Figure S
      key: sfig
      space-before-numbering: false
    - kind: float
      reference-prefix: Table S
      key: stbl
      space-before-numbering: false
---

```{r setup, include=FALSE}
source("dictionaries.R", local = knitr::knit_global())
source("analysisforpaper.R", local = knitr::knit_global())

library(modelsummary)
# source("regularize.R")
```

# Abstract

To gain support, children use signals to communicate their desires to parents. Infant signals of need, particularly infant cries, have been extensively studied in a range of populations. However, the full range of potential child signals of need, which extend beyond cries, has never been investigated in children of all ages. To address this, we used exploratory analyses on caregiver reports of child signaling data from xxx families (xx children) on Utila, Honduras. We found that child signaling decreased with age and neighborhood quality and was highly correlated with the frequency of conflict between children and caretakers. Consistent with signaling theory, children who were perceived as needier within the household were more likely to receive more investment. They were also more likely to receive a pro-child response when the cause of their last instance of signaling involved discomfort, pain, injury, or, illness, but less likely to receive a beneficial response when it involved prior punishment. Our results show there is much potential to the further study of sadness, crying, and temper tantrums among older children using signaling theory.

# Introduction

Compared to chimpanzee mothers, mothers in hunter-gatherer societies have shorter interbirth intervals and produce offspring that require many more years of parental care [@davison2021; @davison2022]. Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@helfrecht2016; @lawson2016; @hagen2009].

## Conflicts of interest within the family

Although the fitness prospects of children and their familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents and other close kin will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974; @bossan2013] because each child shares more genes through descent with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%). Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in their other children, whom parents are expected to value identically, all else being equal [@trivers1974; @bossan2013].

For the same reason, siblings also often find themselves in conflicts of interest with one another, with each being expected to prefer more resources for themself than others. Sibling competition is evident across taxa and can range from relatively benign forms like strong begging competition in chicks [@caro2016] to siblicide in birds, mammals, and sharks [@salmon2014]. In humans, sibling competition is largely estimated by counts of siblings, and large numbers of siblings have been associated with decreased growth [@hadley2011; @hagen2006; @salmon2014], survival [@lawson2012; @meij2009; @strassmann2011], and reproductive success [@gibson2011; @mace1996] in multiple populations, highlighting the likely significance of sibling competition for resources, even if this does not always manifest with behavioral conflict [@bateson1994].

While parent-offspring conflict and sibling competition are likely ubiquitous, the resources being contested will vary across human groups. In relatively egalitarian and food constrained populations, conflict over resources is likely to be centered on food, which probably explains the negative impacts siblings have on child age-specific heights and weights among the Shuar of Ecuador [@hagen2009], although similar effects have also been found in industrialized populations [@lawson2008]. In societies with greater inequality, competition is more likely to be over wealth, as is the case with older brothers in pastoralist communities like the Gabbra and Arsi Oromo limiting the marriage prospects of laterborn sons [@gibson2011; @mace1996]. Compared to large-scale investments like bride price or inheritance, little is understood about smaller transfers of wealth to younger children despite the potential importance of such transfers to 1) facilitate the development of embodied capital outside of growth and fat reserves or 2) help children succeed in their peer groups.

## Cooperation within the family

Humans are a uniquely interdependent species among the living primates [@tomasello2012] in that one’s fitness prospects depend greatly on a network of close kin and other social allies [@apicella2019]. This need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015]. Much of this investment comes from mothers and fathers [@kramer2018]. However, grandparents, aunts, uncles, and non-relatives also provide significant investment in children in many societies [@kramer2018; @sear2008], which was likely favored due to a combination of inclusive fitness benefits, reciprocity, and reputation management [@gurven2000] + (Schniter Preprint (permission needed). Children also often serve as alloparents, helping to raise their younger siblings via direct care, food collection and preparation, and other family tasks [@hrdy2009; @meehan2009; @sear2008] despite still requiring substantial amounts of investment themselves [@kramer2018a].

Children’s potential to invest in their families can also cause conflicts with their parents. Older children might prefer to spend their time investing in romantic or other social relationships, or learning valuable skills, whereas parents might prefer them to care for their siblings, acquire or prepare food, or complete other household chores [@lowe2003; @syme2023a]. Conflicts of interest over investment in the family also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household in order to provide more parental investment to other children [@kushnick2023].

## Signals of needs and wants

In order to elicit more support, children often use signals of needs and wants. Across the tree of life, signals evolve to reveal agents’ private information to other agents when doing so benefits both the signaler and the receiver [@caro2016; @maynard-smith2004; @preuschoft2000; @setchell2005; @zahavi1975]. Children often have needs and wants, where needs involve avoiding potential fitness costs, such as poor nutrition, injuries, or threats to status, and wants involve acquiring potential fitness benefits, such as highly nutritious foods and valuable artifacts. Needs and wants are often not apparent to caregivers, however, i.e., they are private information, and hence children must signal them to receive help. Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, on average, whereas cues are information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995]. Crying when injured is likely a signal of need[\[1\]](#_ftn1), for instance, whereas bleeding would be a cue of need.

When conflicts of interest over investment exist, there may be much to gain from deception (e.g., crying despite the lack of need) so that one could receive the benefits that come from being seen as needy [@zahavi1993]. Unchecked, deceptive signals could result in receivers ignoring all signals.

Credible signals can nevertheless evolve and be maintained in multiple ways. Signals can be trusted in instances when there are no conflicts of interest over what information is conveyed [@guilford1995; @maynardsmith1995]. They can also arise when signals cannot be faked (i.e., indexical signals). Credible signals can also evolve when the benefits of sending a signal outweigh the costs for individuals with the underlying quality or state, but otherwise the costs outweigh the benefits [@spence1973; @zahavi1975; @chisausky2023]. Such signals are often termed “costly”, although what is costly is deception, not necessarily signaling per se [for discussion, see @chisausky2023].

In the context of child signals of need, a child has access to information about their internal state, which is unavailable to others, that allows them to estimate their valuation of additional investment. By signaling in a credible way, the child reveals their resource valuation to potential caretakers who may be skeptical of the child’s claims due to the common conflicts of interest described above. With the information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing. When this is the case, all parties are better off than they would be without the signal, with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response.

However, even credible signals of needs or wants might not result in a beneficial response because belief in a child’s need is not the only factor determining the form and amount of support given by caregivers. Instead, caregivers are expected to weigh the costs and benefits of support against other options [@daly1987; @daly1983]. Ancestrally, this likely often meant evaluating the viability of the child as well as how much a particular resource benefited one child relative to other children, current or future[@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999]. The former is thought to have resulted in children having to demonstrate their vigor early in life [@soltis2004], while the latter is expected to result in context-specific support [@beaulieu2008; @frankenhuis2022]. For example, one unit of a resource is often expected to be worth more to a newborn than an adolescent, favoring that investment flowing to the newborn all else being equal. However, if resources are in short supply, parents may benefit more from investing more in older children as they are closer to reaching adulthood and have higher reproductive value [@daly1983]. Additionally, a parent may also decide against investing in a child, despite deeming that investment worthwhile, if time is not right (e.g., if a mother preferred to breastfeed during rest rather than foraging) [@altmann1980]. When this is the case, outward conflicts may exaggerate the extent of a conflict of interest and involve parents teaching children ways to behave that benefit them both [@altmann1980]. Finally, caregivers often have better information about the costs and benefits of various items and actions, and therefore can make more beneficial decisions for a child than the child can herself (e.g., preventing a child from playing with matches).

Saved for now: Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on their that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].

Saved for now: Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996]. In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971].

### Crying

Crying is the most studied way children signal need. Crying is a human universal [@gracanin2018] with few suspected non-signaling functions outside of the role vocal cries might play in the development of vocalization [@wells2003] and the functions tears have in terms of cleaning, protecting, and lubricating the eye [@murube2009; @sznycer2023]. Hunger, pain, anger, fear, and separation are all common causes of crying [@vingerhoets2013; @wells2003], although tear production also often happens with joyful events [@sznycer2023; @vingerhoets2013; @zickfeld].

Infant cries increase amounts of care, resources, and attention infants receive, as well as increase their proximity to parents [@lin2012]. For example, Karo Batak infants who cried more were breastfed more, and the length of crying bouts predicted the time mothers breastfed [@kushnick2006]. Similarly, !Kung San mothers frequently respond to crying infants by providing comfort and breastfeeding access, with comforting behaviors being far more common than behaviors restricting infants [@kruger2010]. Although this study found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with the majority of the cries ignored lasting under ten seconds. Similar responsiveness has also been reported among the Aka [@hewlett1998], the Mbendjele [@chaudhary2024], the Efe [@tronick1987], and the Ache, among whom young children have been described as “win\[ning\] most conflicts with parents simply by crying and whining” [@kaplan1987]. Studies examining the potential for long-term benefits are rare and difficult to achieve; however, during a time of high infant mortality in a sample of 13 Massai infants, those with easy temperaments were more likely to die (5/7) than those with difficult temperaments (1/6), an outcome that is consistent with signals eliciting beneficial treatment [@devries1987].

Crying does not always succeed in eliciting benefits, however. Compared to the previously mentioned groups, the Ngandu [@hewlett1998] and the Hadza [@blurtonjones1993] have been found to be more likely to ignore the cries of infants and young children, and concerns that being too responsive will spoil children are reasonably common in some Western populations [@barnett2010; @burchinal2010; @chaudhary2024; @frankenhuis2022; @maute2018]. Crying can also lead to negative responses beyond being ignored [@barr1990; @soltis2004; @zeifman2017]. It has been found to increase parental distress, the likelihood of maternal depression, and the child’s risk of maltreatment [@frodi1985; @fujiwara2011; @howard2006; @lin2012; @soltis2004]. For example, in a sample of 3,259 parents of Dutch infants under 6 months old, 5.6% self-reported smothering, slapping, or shaking their babies in response to their crying, outcomes that were more likely among parents who were single, unemployed, or deemed their infant’s crying to be excessive [@reijneveld2004]. Excessive infant crying, i.e., colic, has also been found to be a risk factor for multiple negative outcomes [@zeevenhooven2018], including abuse, shaken baby syndrome [@kurth2011], and earlier cessation of breastfeeding [@howard2006].

Although widely accepted as a signal, there is no consensus on why emotional crying might elicit positive responses. As the costs and benefits of signaling and gaining investment change with age, it is likely the strategic logic of crying changes too. It could be vital for newborns to signal health and vigor to reduce the risk of neglect or infanticide experienced by children with poor survival prospects [@daly1984; @soltis2004]. As crying comes with physiological costs [@rao1997; @thureen1998], it could serve as a credible signal of vigor if these costs are too high for unhealthy infants to bear [@furlow1997; @lummaa1998; @soltis2004; @wells2003]. Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @xx, @yy]. More generally, contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021]. For older children, the relatively insignificant energetic costs of crying might not be enough to ensure it is a credible signal when there are conflicts of interest. For this reason, crying among older children might be more similar to adult crying, which can increase support and empathy but sometimes also reduces reputation and increases stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016]. Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal [@sznycer2023] when there are moderate conflicts of interest.

### Sadness

Sadness is a type of negative affect that is generally thought to arise from loss [@horwitz2017; @levenson1999; @smith1993], especially when one has little power to reverse or mitigate the loss on one’s own [@emanuel2023; @gunderson2021]. Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid future such adverse events [@andrews2009; @nesse1990; @thornhill1989]. Low mood, for example, is often thought to be adaptive in situations when activity is dangerous or unlikely to result in benefits, with anhedonia, pessimism, and fatigue all helping reduce behavior that may not be worth the costs or that may cause additional problems [@nesse2019; @nettle2009]. It may also encourage the reappraisal of one’s circumstances after new information comes to light that one’s current strategy is not working [@nesse2009].

Through the reduction in behavior and attention in one’s usual interests that can accompany sadness and low-mood [@nettle2009], one’s relatives and other social partners would likely gain some indication, in the form of cues, that the sad individual was experiencing adversity and could benefit from increased support. In addition to this cue-based information, sadness also results in facial expressions and crying, both of which probably evolved to indicate need [@balsters2013; @bowlby1980; @reed2017; @darwin1872]. Such expressions seem to be important to conveying this information as is seen by responses to those who cannot produce certain elements of facial expressions like tear production [@hasson2009; @nesse2019] or certain facial movements [@bogart2012; @bogart2010].

### Temper tantrums

Temper tantrums may also be the result of adaptive strategies by which children can elicit more support from parents [@koch2003]. They are common in children from 18 months to 4 years old [@potegal2003] but also exist among older children, albeit with decreased frequency [@macfarlane1962; @barbarin1999; @ward2013; @osterman2010; @goodenough1931; @bhatia1990]. Temper tantrums vary in their severity and include behaviors ranging from crying, shrieking, and screaming to violent bodily movements, aggression, and self-harm [@potegal2003]. Similar behaviors are also seen in other young animals, often surrounding feeding or weaning conflict [@altmann1980; @goodall1986; @schaller1964]. Tantrums have been reported in many societies with different forms of subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1974; @ward1970; @johnson2003; @bhatia1990; @myers1988; @dubois @konner1972], suggesting they might be a human universal [@potegal2019]. As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses, neutral responses (e.g., parents ignoring the child) [@hilger1957; @konner1974; @johnson2003], and negative responses like abuse [@schmitt1987].

As with the vocalizations that come with crying, temper tantrums might motivate caretakers to upregulate their support of the signaler due to the energetic costs of tantrums or the increased risk of predation stemming from the noisy display, both of which might result in credible signals if they are costly enough to deter faking [@lummaa1998]. In addition, tantrums also impose direct costs on caretakers who may be less able to complete important tasks [@altmann1980] like finding or preparing food or taking care of other children.

Another possibility is that temper tantrums carry the risk or implication of self-harm. For example, Utilian mothers reported in focus groups that some temper tantrums involved children banging their heads on things when they do not get what they want, behaviors that have been reported in other populations [@belden2008; @burrows1953; @daniels2012; @dennis1940; @vandenakker2022]. Similarly, young chimpanzees have been observed to beat their heads on the ground when refused access to breast milk [@goodall1986], and American white pelican chicks have been found to throw themselves on the ground, flail violently, and bite at their own wings when begging for food [@schaller1964]. Such behaviors s could credibly signal need as they are likely relatively less costly to those who are truly in dire need compared to those who are not, thus deterring deceptive signaling [@hagen2008].

## Study Aims

Although infant signals of need, particularly infant cries, have been extensively studied in many populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages. In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023a].

This study aimed to establish the natural history of child signaling in Utila, Honduras, by quantifying the frequencies of common child signals of need and various types of caregiver responses. From these, we used signaling theory to explore 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how parents respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses in two subpopulations that face different levels of social and economic adversity. More specifically, we hypothesized that children have evolved strategies to tailor their signals of need to 1) their valuation of a given resource, 2) the presence of existing information about a child's needs and wants (i.e., potential to benefit from a resource) and 3) the extent of conflicts of interest with caretakers, which should influence signal cost. Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.

A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average. This hypothesized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children, especially given the lack of prior studies on signaling by older children in low income settings. For this reason, all analyses of these unique data were exploratory and meant to lay the foundation for future studies of child signaling.

# Methods

## Study population

This study of children aged `{r} paste(range(modeldf$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean. Participants in the study were sampled from two populations who face different levels of economic and social adversity: 1) a long-standing English-speaking community primarily comprising descendants of British immigrants who arrived on the island in the 1800’s, and 2) a more recent community of Spanish-speakers, mostly from the mainland. Compared to those born on Utila, immigrants from the mainland [@davis2014] are more likely to live in crowded and unsanitary conditions, face active discrimination [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000]. Such inequality appears to be increasing in response to recent growth in the tourism industry, with the majority of wealth from tourism being enjoyed by a small portion of native English-speaking islanders and other Westerners [@davis2014].

## Focus group interviews

To inform development of our survey questionnaire on child signaling, focus groups of Utilan parents were held in July 2022. Parents indicated frequent sibling competition and various ways children might signal need. Both Spanish-speaking and English-speaking moms reported that fighting between siblings was extremely common, with fights over toys, gifts, and phone time being the most reported causes. In both populations, crying and various forms of withdrawal were listed as the most common responses to a child not receiving investment or support, followed by temper tantrums and making messes in their home. Spanish-speaking mothers reported that refusing to do homework was a common tactic, while English-speaking mothers emphasized running away and self-harm (e.g., a child slamming their head on the ground) to a greater degree. In response to these hypothesized signals, discipline was common. Although the mothers generally reported that children were treated equally, needier children were described as getting more resources from parents, perhaps due to differences in signaling among children as well as information caretakers gained from cues of need.

## Sampling

Participants for this study were recruited through convenience sampling. JH recruited early participants with the help of primary informants and the use of snowball sampling. Participants were also recruited via Facebook posts and fliers which were sent home with children attending Utila's Methodist school.

In all forms of recruitment, adult participants were informed of the purpose of the study, what they would be asked to do, and the inclusion criteria, such as being 18 or older, having permanent residence on Utila, and having a child between 6 and 18 years old.

\[no longer needed\] XXXXXXXXCHIld RecruitmentXXXXXXXXXXX

## Participants

All participants for the survey portion of the study met the inclusion criteria. Most ~~adult~~ participants were the biological mothers of one or more children in the study. ~~All children in the study, whose ages ranged from 5-20, were cared for by one of the adult participants.~~

## Ethics

All ~~adult~~ participants provided informed consent. The surveys, which included many topics outside of child signaling, were designed to last about an hour, plus 20 minutes for each additional child. ~~Adults~~ Participants were paid \$XXXXX for the survey portion of the study.

~~xxxxxxxxxxxxxChild consent infoxxxxxxxxxxxxxx.~~

This study was approved by the xxxxxxxIRBINFOxxxxxxxxxxxxx.

## Survey

The data for this study were collected as part of the Utila Child Health Project. Participants could choose to do the survey in English or Spanish. All questions were read to them and recorded by a research team member.

### Frequency of child signaling, alloparenting contributions, and conflict with caretaker

We used 6-point Likert scales to get caregiver reports on the frequencies of 1) various types of child signaling, such as sadness and crying, 2) conflict between the caretaker and each participating child, 3) and child allocare of other children in their household. See @tbl-freqvars for exact wording. Caretakers were often not asked about child alloparenting behavior if a child was an only child. For this reason, we gave children an alloparenting frequency of “never” if they were the only child within a household.

|  |  |
|-------------------------|-----------------------------------------------|
| Variable | Survey question wording |
| Crying Frequency | How often does your child cry? |
| Sad Frequency | How often is your child sad? |
| Tantrum Frequency | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency | How often does your child leave the house without telling you when they get mad? |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child? |
| Child Alloparenting Frequency | How often does your child take care of their other siblings? |

: Variable names and question wording for measures about the frequency of signaling, allocare, and conflict. All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day. For analyses, we converted these categorical variables to numeric frequencies per month as follows: never = 0; once a month or less = 1; more than once a month = 3; more than once a week = 8; daily = 30; multiple times per day = 60. {#tbl-freqvars}

From these count variables we created three composite signaling variables to be used as outcome variables. The *frequency of signaling* was the sum of the frequencies of child sadness, crying, and temper tantrums. *Maximum signal frequency* was the frequency of the most common of these three signals. *Signal cost* was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., *Signal cost* = *frequency of sadness* + 2 $\times$ *frequency of crying* + 3 $\times$ *frequency of temper tantrums*).

We also calculated the summed frequency of child alloparenting contributions by other children within the household.

### Relative neediness of children and relative investment in children within the family

We used 6-point Likert scales for parental reports of 1) the relative neediness of children within the household, and 2) the relative investment they provide to each participating child compared to the rest of their children. Both questions were framed need as involving time, effort, and resources. See Table @stbl-relvars for exact wording.

### Causes of, and responses to, the last instance of child signaling of need

We asked caretakers about the last time each participating child was sad or cried. Caretakers provided a free response of the reason(s) for the emotional display, and their responses to it. The researcher then binned responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective). For the full list of options see @stbl-response-valence. We then created a *caregiver response* variable, which was coded as 1 for a positive response (e.g., helping child), 0 for a neutral response (e.g., not responding), and -1 for a negative response (e.g., punishment; all from the child's perspective). MG and KS also coded the causes for the presence or absence of 17 elements (e.g., whether the caused involved the child was asked to do an *unwanted task*, *punishment*, *family conflict*, and/or *discomfort, pain, injury, or illness*).

### Caretaker perceptions of child health and ability to avoid illness

Caretakers were asked about the frequency of their child missing school and their child's history of medical problems in general terms (i.e., the extent to which they experienced a "variety of medical issues"). Both involved the use of five point Likert scales, the scores of which were averaged for use in analyses. See @stbl-medical-record for the wording of these variables.

Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness. We reversed the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean *illness avoidance* variable. In the `{r} sum((!is.na(modeldf$IllnessSusceptibility2)) & is.na(modeldf$IllnessSusceptibility3))` instance in which we only had data for 2 of the measures, we still calculated the mean but used a numerator of 2. See @stbl-healthvars for the wording of these variables.

### Caretaker perceptions of house quality, neighborhood quality, home instability, and food security

Affirmative caregiver responses to the following questions about their home were summed to create a home quality index: having an indoor bathroom, having a bathroom that is not shared with anyone outside of the household, having a flush toilet, having running water to wash hands in the bathroom, and having a home which did not flood in the last year.

Caregiver perceptions of neighborhood quality were measured using ten questions relating to the costs and benefits associated with living in one's neighborhood. The mean of these values was used in analyses as a neighborhood quality index. Levels of home instability were measured with four questions covering the degree to which the caregiver's family had/has to move homes and whether people move in and out of their homes frequently. For wording of the these items, which used five point Likert scales, see @stbl-Neighborhood-Quality and @stbl-Home-Instability.

We used a modified version of the USDA Food Security Questionnaire [@blumberg1999] to measure food insecurity, with the most significant changes being converting of the measures to a five point Likert scale and averaging them for a food insecurity index. See @stbl-Food-Security for the wording of these variables.

### Demographic questions

The survey also included many demographic questions about the caretakers and their households. Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) kinship ties within households, (4) partnership status of the primary caregiver, (5) number of adults who provide childcare within the household, (6) highest level of education of the primary caregiver, (7) immigration status (in regards to Utila), (8) monthly household income, (9) neighborhood of residence, (10) current job of the primary caretaker, and (11) ownership of various possessions (e.g., homes, washing machines, or cellphones; see @stbl-possession-list for full list). These data allowed us to calculate the number of children (total, younger, and older) and adults within the household as well as the genetic relatedness between siblings. We also created a variable for residence in Camponado, a densely populated neighborhood on the island with relatively poor sanitation in which most individuals from the mainland live that is seen as the most dangerous and poorest part of Utila [@garcia2017].

In one instance where data for caregiver age was missing, we set the missing value to the mean. This affected data for two children.

## Statistical analyses

All statistical analyses were exploratory. Whereas confirmatory studies test prespecified hypotheses generated blind to the data, exploratory analyses are data-driven and prioritize discovery: they risk overfitting the data, generating false positives, but can reveal unexpected yet potentially important relationships that warrant future investigation [@tukey1977].

The overall structure of our data was analyzed with principle components analysis (PCA), Bayesian graphical models (BGM), and minimum spanning trees (MST). PCA is a linear transformation of the data to a new orthonormal coordinate system in which the the first principle component (PC1) corresponds to the axis of maximum variance, the second principle component (PC2) corresponds to the axis perpendicular to the first of the maximum of the remaining variance, and so forth. Among other uses, PCA identifies the linear combinations of variables that account for most of the variance, thereby providing a lower dimensional representation of high dimensional data. In BGA, variables are nodes in an undirected graph, and edges represent their conditional dependencies. Specifically, an edge indicates that two variables are dependent, conditional on the other variables; edge weights indicate the strength and sign of the relationship. Importantly, the absence of an edge indicates that two variables are conditionally independent [@huth2023]. The graph can therefore identify clusters of dependent variables that are independent of other variables. From the graphical models identified with BGM, we extracted an MST, i.e., a subset of the graph in which each node (variable) is connected without any cycles and with the minimum total "distance" between the nodes (more strongly dependent variables, i.e., those with larger absolute weights, having lower "distances"). The MST is therefore typically a dramatically simplified representation of the original graph that often identifies the most important edges among the nodes, which in our case are the most important dependencies among the variables.

To investigate the associations between potential causes of child signaling and conflict with caretakers and caregiver reports of child behavior, we used elasticnet regression models. Elasticnet models are useful in cases where the number of predictor variables (p) is large relative to the number of observations (n). These models, often referred to as penalized or regularized regression (e.g., lasso and ridge regression), aim to limit overfitting the data (variance) by allowing some bias, i.e., by optimizing a bias-variance tradeoff. In particular, lasso regression sets the coefficients of uninformative variables to exactly 0, thereby serving as a form of variable selection [@zou2005].

Elasticnet regression models of signaling and conflict frequency, which were overdispered count data, were fit using the glmnet package [@friedman2010; @wurm2021] and the quasipoisson error family. Predictors included demographic, household, health, and neighborhood variables that might cause signaling behavior or conflict between children and caretakers. All variables were centered and standardized by one standard deviation prior to fitting.

Parental responses to child signaling, their perceptions of child need, and their reported investment in the child were coded as three-level ordinal variables, either negative, neutral or positive; or less than other children, same as other children, or more than other children. We analyzed these outcomes using ordinal logistic regression with elastic net penalties, fit with the ordinalNet package [@wurm2021].

All analyses were conducted with `{r} R.version.string`.

## Data availability

The anonymized data used in the study are available at https://github.com/grasshoppermouse/utiladata2023. Study code is available at https://github.com/michaelrgaffney/utilasignaling.

# Results

`{r} dplyr::n_distinct(modeldf$householdID)` primary caretakers (females = `{r} table(caregivers$CaregiverSex)[1]`, males = `{r} table(caregivers$CaregiverSex)[2]`, non-binary = `{r} table(caregivers$CaregiverSex)[3]`) took the survey and provided signaling data for at least one child. `{r} nrow(modeldf)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three. Among those with at least one record of signaling frequency, `{r} sex_table[1]` were female and `{r} sex_table[2]` were male. Anthropometric and signaling data exist for `{r} nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `{r} sex_table2[1]` were female and `{r} sex_table2[2]` were male. For summary statistics of study variables see @tbl-sumtab_combined and @tbl-sumtab_factors.

```{r sumtab_combined}
#| label: tbl-sumtab_combined
#| tbl-cap: Summary statistics for continuous variables.

kable(custom.summarize(modeldf, longform_dict[1:23])) %>% kable_styling()
```

```{r sumtab_factors}
#| label: tbl-sumtab_factors
#| tbl-cap: Summary statistics for factor variables.

# convert to factor to omit NAs in datasummary_skim
dx <- modeldf |>
  dplyr::select(any_of(names(longform_dict))) |>
  mutate(
    across(starts_with(c("Lifestyle", "ImmigrateUtila")), as.factor)
  )

names(dx) <- longform_dict[names(dx)]
datasummary_skim(data = dx, type = "categorical")
```

A PCA of primary study variables found that sociodemographic and ecological variables loaded primarily on PC1, whereas the child signaling, conflict, and child age variables loaded primarily on PC2. Child signaling variables modestly load on PC1 as well in a direction suggesting that greater signaling covaries with more adverse environments. See @fig-PCA.

::: {#fig-PCA}
![](Figures/plot_pca.svg)

A)  Loadings of PC1 and PC2 on each variable. B) Biplot map of PC1 and PC2. Each point is one child. Arrows represent the loadings of PC1 and PC2 on each variable.
:::

We then estimated a BGM for all study variables, extracting the MST, which revealed that signaling variables clustered together, conflict was positively associated with signal cost, and child age negatively associated with crying frequency. See @fig-mst.

::: {#fig-mst}
![](Figures/plot_mst_1e6_cost_freq.edited.png)

Minimum spanning tree of child signaling and variables thought to cause differences in signaling strategies. Blue edges = negative \[do we want more detail here? e.g., partial correlations or w/e the proper Bayesian way to word that is\] correlations between variables. Red edges = positive correlations between variables. Edge lengths represent \[check that this is right and if more specificity is needed\] distances between variables. Line weights represent(1 - correlation coefficient)
:::

## Distributions of signaling behavior, alloparenting effort, and conflict

```{r}
fmt_sum <- function(v){
  v <- na.omit(v)
  n <- length(v)
  thesum <- sum(v)
  thepct <- 100*signif(thesum/n, 2)
  str_glue("{thesum} children ({thepct}%)")
}
```

Child sadness, crying, and temper tantrums were common and highly correlated with one another. See @fig-barplot_SignalFreq and @fig-signal_corrplot. Caregivers reported that `{r} fmt_sum(modeldf$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(modeldf$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(modeldf$TantrumFreqN >= 8)` threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(modeldf$SadFreqN >= 30)`; crying: `{r} fmt_sum(modeldf$CryFreqN >= 30)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN >= 30)`. However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(modeldf$SadFreqN == 0)`; crying; `{r} fmt_sum(modeldf$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN == 0)`. For a visualization of how signal frequencies varied within children see @sfig-signal_alluvial_plot.

Conflict between children and caretakers was also common, with `{r} fmt_sum(modeldf$ConflictFreqN >= 8)` having conflicts with parents more than once a week. However, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents. See @fig-barplot_ConflictFreq. Conflict was also associated with higher signaling frequencies. See @fig-PCA and @fig-signal_corrplot.

The frequency of child alloparenting effort, which previous research has linked to parent-child conflict, was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day. See @fig-barplot_AlloparentingFreq.

::: {#fig-barplot_SignalFreq}
```{r}
barplot_SignalFreq + guides(fill = guide_legend(title = 'Times/month', nrow = 1, position = 'top', reverse = T))
```

Frequencies of child sadness, crying, and temper tantrums per month as reported by the primary caretaker. Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
:::

::: {#fig-signal_corrplot}
```{r}
signal_corrplot
```

Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers. For a more complete correlation matrix of study variables see @sfig-plot-full-corrmat.
:::

::: {#fig-barplot_ConflictFreq}
```{r}
barplot_conflict
```

Frequency of conflict (per month) between child and primary caregiver as reported by primary caretaker. Color corresponds to sex of the child. Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
:::

::: {#fig-barplot_AlloparentingFreq}
```{r}
barplot_alloparenting
```

Frequency of a child's alloparenting effort (per month) as reported by primary caretaker. Color corresponds to sex of the child. Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
:::

## Predictors of signaling frequency and conflict frequency

Using lasso regression, we explored all potential predictors of each child signal frequency, as well as of conflict frequency. See @fig-lasso.

::: {#fig-lasso}
![](Figures/plot_lasso.svg)

Lasso regression models of each child signaling frequency and conflict frequency. Variables whose coefficient values were shrunk to 0 are not displayed. Models fit using quasi-Poisson regression.
:::

The effects of particularly large risk or protective factors for child signaling are shown in @fig-signal-effects and @fig-alloparentingXsex.

::: {#fig-signal-effects}
![](Figures/signal_effects_plot.svg)

Effects of selected predictors of child signaling frequencies when holding continuous variables at their mean and factor variables at their mode.
:::

::: {#fig-alloparentingXsex}
![](Figures/plot_alloparentingXsex.svg)

Effects of alloparenting frequency (per month) and signal cost and frequency of temper tantrums by sex.
:::

The effects of particularly large risk or protective factors for conflict frequency are shown in @fig-conflict-effects.

::: {#fig-conflict-effects}
![](Figures/conflict_effects_plot.svg)

Effects of selected predictors of child-parent conflict frequencies when holding continuous variables at their mean and factor variables at their mode.
:::

## Positive associations between child age and signaling frequency and perceptions of child neediness

Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children. Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor. The end result was `{r} rel_need_table[1]` less needy children, `{r} rel_need_table[2]` children who were as needy as their siblings, and `{r} rel_need_table[3]` more needy children.

There were positive relationships between perceived relative neediness within the family and both child age and the frequency of child sadness. See @fig-effect_relativeneed.

::: {#fig-effect_relativeneed}
![](Figures/plot_need_combined.svg)

The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children. A: Effect of child age on relative neediness B: Effect child sadness frequency on relative neediness. Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
:::

## Relative need predicts relative investment

Signals are capable of eliciting better treatment because the information they provide alters the behavior of signal targets in ways that benefit the signaler. For this reason, we tested the association between caretaker perceptions of relative child need within the household and the caretaker's relative investment within the household. Consistent with signaling theory, our model predicts a greater proportion of caretakers investing more in children who they deem needier and a greater proportion of caretakers investing less in children they deem less needy, controlling for the age of the child, and household demographics. See @fig-plot_relative_investment.

::: {#fig-plot_relative_investment}
![](Figures/plot_invest_need.svg)

The effect of perceived need (in terms of time, effort, and resources) relative investment within the household. Outcome variable question wording: How much time, effort, and resources do you put into this child compared to your other children?
:::

## Caregiver responses to most recent signal of need

```{r}
# \[do we want to mention distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NegativeResponseBinary > 0)` negative response). However, they were also more likely to respond in a neutral or negative manner rather than a positive one when caretakers believed signaling came in response to the child being punished (`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$PositiveResponseBinary > 0)` positive responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NeutralResponseBinary > 0)` neutral responses vs. `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NegativeResponseBinary > 0)` negative responses).
```

::: {#fig-effect_last_signal_all_responses}
```{r}
barplot_CaregiverResponse
```

Frequencies of caregiver response to the last instance of child signaling by the cause of the signaling. Categories of causes were based on cluster analysis of more fine-grained cause variables. See sixxxxxx for cluster and sitabxxxxxxx for the frequencies of causes and caregiver responses.
:::

There were `{r} response_overlap[1,1]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective. These were excluded from analyses due to their rarity.

Our models found no evidence of relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective. However, primary caretakers were more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness, and less likely to respond positively when they attributed the cause of signaling to child punishment (any statistic report here?). See @fig-effect_last_signal_response.

::: {#fig-effect_last_signal_response}
![](Figures/plot_caregiver_response_combined.svg)

Effects of selected predictors of the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried. 1 = evidence of the cause in the caregiver report; 0 = no evidence. The endpoints of the lines were dodged slightly for greater visibility. Dot size represents xxxxxxx. See @sfig-pot-caregiver-punish for the effect of the cause involving punishment.
:::

# Discussion

Caregivers reported a wide range of child signaling frequencies. About half of children were reported to cry or throw tantrums more than once a month and almost 40% were reported to do so more than once a week [@fig-barplot_SignalFreq]. Frequencies of sadness, crying, and tantrums were positively correlated with each other and with ~~family conflict~~ the frequency of conflict with the primary caretaker [@fig-signal_corrplot]. They also loaded in the same direction on PC1 as indices of adversity, such as food insecurity, and in the opposite direction of indices of wealth and environmental quality, such as possessions, neighborhood quality, and income [@fig-PCA]. Signaling decreased with child age, and increased with ~~family~~ conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST [@fig-mst]. In our lasso regression models (which excluded family conflict due to the likely bidirectional relationship between signaling and perceived conflict), child age and high neighborhood quality were consistently negatively associated with signaling frequency, and household income, caregiver education, and number of adults helping with childcare generally were too; medical problems were a consistent, albeit small positive predictor [@fig-lasso; @fig-signal-effects].

These results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, and that children's need is higher in adverse environments and at younger ages. One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], is that as children mature, they are more capable of fulfilling their own desires, more able to ask others for what they want [@osterman2010], and less vulnerable to threats. The effects of age on signaling are also likely due to how age influences the costliness of signaling need. For example, while the energetic cost of crying might be costly for the youngest children, it is likely insignificant for older children. Older children might instead need to weigh the reputational harm from crying against the potential increase in social support.

Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, caregiver perceptions of conflict and child signaling frequency were correlated [@fig-signal_corrplot], and family conflict was the most common cause of a child's recent signaling episode [@fig-effect_last_signal_all_responses]. Although conflicts of interest do not always lead to overt behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest. Conflicts, in turn, were negatively associated with child age and higher neigborhood quality, and positively associated with medical problems and (oddly) caregiver education [@fig-conflict-effects]. Finally, more frequent sadness was positively associated with perceived higher need relative to other children [@fig-effect_relativeneed], and higher perceived need was associated with higher relative investment [@fig-plot_relative_investment]. Interestingly, older children were percieved to be needier than younger ones. These results generally support the hypothesis that the signals we investigated are costly, credible signals of need.

\<!- If this is the case, the positive correlation between signaling effort and conflict frequency may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums). It is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself. -\>

\<!- In times when an older child's valuation of potential increased support is small, such costs may be prohibitive. However, in times when a child places more value on potential support, those same costs could allow for a child to credibly signal need if they are too high to be worth it for deceptive signaling but still made up for by the benefits for credible signalers. -\>

Caregiver responses to signals were generally positive or neutral, although less so for situations involving family conflict or transgressions [@fig-effect_last_signal_all_responses; @fig-effect_last_signal_response].

Finally, based on previous research showing that family demands, such as caring for younger siblings, was an important source of parent-offspring conflict [@syme2019], we investigated the relationship between the frequency of alloparenting by children and their frequency of signaling. We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, but this relationship was slightly negative for boys [@fig-alloparentingXsex]. In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers. If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in relationships and status seeking behaviors outside the household.

--- Hagen stopped here ---

\<!- We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys. -\>

Evidence for conflicts similar to the one hypothesized here has been reported by multiple researchers working with the Chuukese. Lowe found that both male and female children reported conflict with their families over their desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003]. In an analysis of conflict among Chuukese-Americans, Syme and Hagen found that severe conflicts were more likely with greater numbers of siblings and when conflicts involved a child spending time away from home or evading family responsibilities [@syme2023a], findings which would be consistent with parents pressuring older children into investing more in the household among other explanations. In response to such conflicts, participants reported using a wide range strategies as children including, deception, withholding cooperation, running away, and seeking the assistance of others. Through these means, participants were often able to resolve conflicts in their favor despite their lower status in the household and community.

However, child alloparenting effort was negatively associated with caregiver reports of child conflict. One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual. In line with this reasoning, we found evidence of a child alloparenting effort by child age interaction, with our model predicting increased levels of conflict for older children with greater alloparenting \[if we say this we should report the results of that lasso too\]. Ideally, we could control for signaling behavior to test this. However, the bidirectional relationship between child signaling and conflict with caretakers prevents this. It is also possible that our hypothesis about conflict over investment in the household is incorrect.

There was a positive association between the frequency of child sadness and caretaker perceptions of child neediness within the family. However, the other signaling measures examined had coefficients set to 0 in the same lasso model despite the expectation that their greater costliness would make them more credible sources of information to parents. One reason for this might be that costs of sadness, while minor short-term, can be significant long-term. These costs are clearly seen in depression, which often involves an extended period of a lack of interest in one's normal activities and decreased investment in oneself and one's cooperative relationships and has been suggested to capable of credibly signal need in times of substantial conflicts of interest [@watson2002; @hagen2003].

Another possibility is that parents are using cues from sadness (i.e., withdrawal and rumination) to evaluate the relative neediness of children. As cues are not created for the purpose of communication, they are completely reliable if perceived and interpreted correctly. Therefore, if some cues are hard to or are impossible to fake, they may be more reliable than most signals. That said, it is likely that some symptoms of sadness are simultaneously the result of both signals and cues, something which may have been favored by natural selection if the threat of deception meant cue-level symptoms were insufficient in eliciting support without upregulation, something which has been suggested to be the case with various illness symptoms [@tiokhin2016; @steinkopf2015].

Caretakers perceived older children as needier, despite having more conflicts with younger children. One likely reason for this might be that caretakers percieve the types of investment older children seek as more costly. In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often during surveys. It may also be that caretakers see investing in older children as relatively costly as they seek to withdraw investment in order to invest in younger children.

Signaling behavior did not predict differences in relative caregiver investment between children within households. However, greater perceptions of need among caregivers were associated with increased likelihood of a child being reported to receive more investment in terms of time, effort, and resources relative to other children within the household. As previously mentioned, credible signals are capable of leading to pro-signaler responses due to their ability to reveal private information to others who would be motivated to increase their investment in another should they have that information. In other words, much of the effect of signaling on levels of support is expected to be mediated by the change in the beliefs of others that one could benefit from increased support that it causes (e.g., that a child is truly in need due to adversity) rather than direct effects. In line with this reasoning, an experimental vignette study which found that depression and suicidality increased the likelihood of costly support from participants in times of uncertainty and substantial conflicts of interest also found that between 67% and 78% of the effect of signal type on the likelihood of costly support was mediated by the effect of signaling on belief in need [@gaffney2022].

Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not. One potential causal explanation for this is that parents may often have information for sources outside of child signaling to corroborate a child's need (e.g., a fever, a diagnosis, or witness a child hurt themselves) so that there is less private information about a child's current state. Simultaneously, conflicts of interest may be less significant in these times due to either 1) both sides benefiting similarly from their child returning to normal (e.g., with an illness) or 2) the cause involving something which is minor and would quickly resolve itself (e.g. a skinned knee). It may also be that many of the typical positive responses to pain and illness are not very costly to caretakers (e.g., providing comfort), thereby increasing their frequency.

When the cause involved the child having made a transgression, punishment, or the loss of privleges or an item, however, parents were more likely to respond negatively and less likely to respond positively. In these circumstances, children may be signaling about the severity of a punishment [@syme2019]. However, parents may be reluctant to reduce punishments if the behavior that warranted the punishment was repeated or severe. They may also react with scolding or further punishment. That said, our model predicts positive responses in `r #signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.

Parents were also more likely to respond negatively and less likely to respond positively to children when the cause involved family conflict. This might reflect tradeoffs within the family, whereby investment in one child cannot go to another. Caretakers may have also responded more negatively when the cause involved family conflict, if it involved a child imposing costs on other family members.

\[delete this paragraph?\] We found a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status. One possible reason for this may be that more educated caretakers see their children less during the day due to greater ability to afford childcare or due to them having jobs that are less compatible with child care (test model of social support predicted by education). Unfortunately, we do not have data on how much time caretakers spent with children to test this, and it seems equally likely that less educated parents may see their kids less often if they need to work more hours to support their families. Another possibility is that lower levels of education are associated with greater levels of adversity for both parents and children, something which appears to be the case for other measures of SES [@pepper2017] and would increase the benefits of signaling need for children and the difficulty of parents to respond. It is also possible that education influences parenting norms in ways that cause differences in child signaling behavior.

Outside of the sex by alloparenting effort interaction, we found little evidence of sex differences in child signaling behavior. As adults, females cry more than males, a finding that has been reported across cultures [@vanhemert2011; @becht2001]. One reason why our study may have failed to find significant effects of sex is that sex differences in the causes and reputational effects of signaling need are small or nonexistent until adolescence and young adulthood, with our small sample size for this age range preventing us from detecting a real effect should it exist. It is also possible that the sex difference seen in rates of crying in adults may reflect the lesser ability of females to bargain for better treatment through aggression or the threat of it when faced with adversity [@hagen2016]. Such a sex difference may not exist among preadolescencents due to a lack of difference in strength or athletic ability before puberty [@handelsman2018] or due to their relative lack of formidability compared to adults.

Despite the differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000] between recent immigrants from the mainland and native Utilians, primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions. However, neighborhood quality, which was strongly negatively correlated with residence in Camponado, (check others) was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients. Monthly income was also negatively associated with crying frequency,\~\~ tantrum frequency (too small of a coefficient to mention here?),\~\~ summed signal frequency, and signal cost albeit with smaller coefficients.

Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000] and multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]]. In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children they may be competing with for status and desirable cooperative relationships. ~~In addition to overall effort in eliciting need, resource availability may have also affect how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].~~

(temporarilly moved here) Tantrums were present among older children and might be an under-appreciated way by which children signal for increased support. Most studies of tantrums involve children under 4 years old or older children with disruptive mood dysregulation disorder, likely due to their greater prevalence in these populations (cite). However, multiple studies which included older children from non-clinical populations have found similar evidence of tantrums [@macfarlane1962; @barbarin1999; @ward2013; @osterman2010; @goodenough1931; @bhatia1990], as have ethnographic studies [@ward2013; @lessa1950; @dennis1940].

## Limitations

The main limitation of this study is that it is completely exploratory, something which results in a high risk of our models overfitting the data and indicating associations which might not exist in the real world. Our study was also predominantly cross-sectional, further limiting our ability to make causal inferences.

Our use of convenience and snowball sampling is also a limitation as it resulted in a non-probability sample which may not be representative of the broader population. This could happen if there was selection bias so that those who participated differed from the rest of the population in how they interacted with their children or how they assessed, remembered, and reported child behavior.

~~Outside of child anthropometric measurements,~~ All data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever they may be. Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we also faced a tradeoff between survey time and question depth. Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, questions focused on specific questions. Framing the question this way may have also increased the likelihood of social desirability bias [@grimm2010], which could lead adult participants to be be more likely to report equal investment across children. Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.

# Conclusion

Children face the challenge of eliciting support from others in times of adversity and opportunity. The signaling strategies of Utilian children were highly correlated and engaged in by children of all ages. Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, 1) child signaling effort was positively correlated with the frequency of conflict with caretakers and 2) greater perceptions of child need predicted greater investment in a given child. Our results suggest there is much potential for signaling theory to help explain patterns of child crying and temper tantrums.

**Footnotes**

[\[1\]](#_ftnref1) We expect “signals of need” to actually be signals of resource valuation. However, we use “signal of need” here to distinguish signals that involve attempts at eliciting support through revealing private information about one’s vulnerabilities from the more commonly studied signals which reveal private information about one’s positive qualities (i.e., those related to partner choice and intrasexual competition).

[\[2\]](#_ftnref2) Our use of “credible” signaling is largely synonymous with the more common term, honest signaling. However, we prefer the term credible signals since signals do not need to be completely honest for natural selection to have favored individuals who paid attention to them.

[\[3\]](#_ftnref3) Researchers often distinguish between sadness and low mood based on the length of negative affect. We remain agnostic on how different sadness and low mood are from one another but use the terms researchers used when describing their research.

[\[4\]](#_ftnref4) All else is often not equal when it comes to the costliness of investment as older children may often cost more (e.g., the costs education and food are often higher among older children). However, parents would still have to pay similar costs to younger children as they grew, something which natural selection has likely acted on when it comes to our adaptations involved with differential child investment.

[\[5\]](#_ftnref5) Even in instances where conflicts begin due to mistakes, the initial conflict can lead to increased conflicts of interest. This may be seen with findings that individuals who accidentally harm another often grow to dislike them (cite), something which has been suggested to be adaptive to the accidental perpetrator as the unintended victim is likely to value than they would have before, potentially to the point of wishing them harm.

[\[6\]](#_ftnref6) Despite the risk, Shostak notes there were no known recent incidents of harm to women who ran away to spend the night in the Bush [@shostak1981].

# Supplementary Information

## Additional information for survey measures

::: {#stbl-relvars}
|  |  |
|-------------------------|----------------------------------------------|
| Variable | Survey question wording |
| Relative Need | How much time, effort, and resources does this child need compared to your other children? |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

Variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.
:::

::: {#stbl-medical-record}
|  |  |
|-----------------------|------------------------------------------------|
| Variable | Survey question wording |
| SchoolAges1 | My child has missed a lot of school due to illness. |
| SchoolAges2 | My child has had to deal with a variety of medical issues in their life so far. |

Variable names and question wording for measures about a child's health history. Both questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree."
:::

::: {#stbl-healthvars}
|  |  |
|--------------------------|----------------------------------------------|
| Variable | Survey question wording |
| IllnessAvoidance1 | If an illness is 'going around', my child will get it. |
| IllnessAvoidance2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| IllnessAvoidance3 | My child’s immune system protects him/her from most illnesses that other kids get. |

Variable names and question wording for measures about a child's current likelihood of resisting illness. All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for IllnessAvoidance1 and IllnessAvoidance2 were inverted so that 5 corresponded with the lowest susceptibility to illness, i.e., greatest *illness avoidance*.
:::

::: {#stbl-Food-Security}
|  |  |
|-----------------------|------------------------------------------------|
| Variable | Survey question wording |
| FoodSecurity_1 | In the past year, the food that we bought often didn’t last, and we didn’t have money to get more. |
| FoodSecurity_2 | In the past year, we couldn’t afford to eat balanced meals. |
| FoodSecurity_3 | In the past year, I (or other adults in my household) sometimes cut the size of my meals or skipped meals because there wasn't enough money for food. |
| FoodSecurity_4 | In the past year, I sometimes ate less than I felt I should because there wasn't enough money for food. |
| FoodSecurity_5 | In the past year, I was sometimes hungry but didn’t eat because there wasn’t enough money for food. |
| FoodSecurity_6 | In the past year, I ate lower quality food because I couldn't afford better food. |

Variable names and question wording for the modified version of the USDA Food Security Questionnaire [@blumberg1999]. All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.”
:::

::: {#stbl-Home-Instability}
|  |  |
|------------------------|------------------------------------------------|
| Variable | Survey question wording |
| HomeInstability_1 | We move our home a lot. |
| HomeInstability_2 | I wish our household felt more stable. |
| HomeInstability_3 | I worry that we will be forced to move before we are ready. |
| HomeInstability_4 | People move in and move out of our home frequently. |

Variable names and question wording for the home instability questions. All questions were asked using a Likert scale from 1-5 where 1 was labeled “Disagree strongly” and 5 was labeled “Agree Strongly.”
:::

::: {#stbl-Neighborhood-Quality}
|  |  |
|--------------------------|----------------------------------------------|
| Variable | Survey question wording |
| NeighborhoodQuality_1 | As a place to live, my neighborhood is good quality. |
| NeighborhoodQuality_2 | As a place to raise children, my neighborhood is good quality. |
| NeighborhoodQuality_3 | My neighborhood quality has gotten worse over time. |
| NeighborhoodQuality_4 | My neighborhood is more dangerous than other parts of the island. |
| NeighborhoodQuality_5 | Sometimes I hear about a fight with a weapon in my neighborhood. |
| NeighborhoodQuality_6 | Sometimes I hear about illegal drug activity in my neighborhood. |
| NeighborhoodQuality_7 | Sometimes I hear about someone being badly hurt in my neighborhood. |
| NeighborhoodQuality_8 | Sometimes I hear about someone's home being robbed in my neighborhood. |
| NeighborhoodQuality_9 | Sometimes I hear about someone being mugged in my neighborhood. |
| NeighborhoodQuality_10 | Sometimes I hear about someone being assaulted in my neighborhood. |

Variable names and question wording for the perceptions of neighborhood quality questions. All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.”
:::

::: {#stbl-response-valence}
Positive: Give resources (e.g., toys, food, money)\
Positive: Promise resources\
Positive: Give attention/affection\
Positive: Promise attention/affection\
Positive: Give opportunities (e.g., loosen restrictions, new privileges, time with friends, fun activities)\
Positive: Promise opportunities\
Positive: Remove responsibilities (e.g., fewer chores)\
Positive: Promise removal of responsibilities\
Positive: Take child's side in conflict (e.g., taking child's side against sibling, intervening against bully)\
Positive: Give advice\
Positive: Other \[Text Box\]\
Neutral: Distract\
Neutral: Do nothing\
Neutral: Explain (e.g., go over why the child cannot have or do something)\
Neutral: Other \[Text Box\]\
Negative: Threaten removal of resources\
Negative: Remove resources (e.g., toys, food, money)\
Negative: Threaten withholding of attention/affection\
Negative: Withhold attention/affection (e.g., silent treatment, ignoring)\
Negative: Threaten loss of opportunities (e.g., freedom to move around the island, beach time, timeout, time with friends)\
Negative: Loss of opportunities\
Negative: Threaten more responsibilities (e.g., chores, alloparenting, homework)\
Negative: Add responsibilities\
Negative: Support adversary (e.g., taking sibling's side)\
Negative: Threaten physical punishment\
Negative: Punish physically\
Negative: Punish verbally (includes yelling)\
Negative: Downplay adversity\
Negative: Other \[Text Box\]\
NA (Can't remember, refuse to answer, etc)\
Other: Other \[Text Box\]

List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried.
:::

::: {#stbl-possession-list}
Motorized vehicle (scooter, 4-wheeler, etc.) Bicycle Television (TV) with cable Smartphone Washing machine Refrigerator A home that you own Credit card

List of possessions owned (yes or no).
:::

## Figures

::: {#sfig-signal_alluvial_plot}
```{r}
signal_alluvial_plot
```

Combinations of signaling frequencies per month. Alluvium width equals the number of individuals with the same frequencies for the three signals/states. Stratum height equals the count of individuals at a given frequency. Frequency key: Low includes "never" and "once a month or less," medium includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and "multiple times per day."
:::

::: {#sfig-plot-full-corrmat}
![](Figures/plot_fullcorrmat.svg)

Complete correlation matrix of study variables.
:::

::: {#sfig-pot-caregiver-punish}
![](Figures/plot_caregiver_punish.svg)

The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried when punishment was implicated in the cause (0 = no punishment, 1 = punishment.)
:::

## Coefficient plots

::: {#sfig-plot-need-coefs}
![](Figures/plot_need_coefs.svg)

Effects of predictor variables on relative need between household children. Ordinal logistic regression model with elastic net penalties.
:::

::: {#sfig-plot-invest-coefs}
![](Figures/plot_invest_coefs.svg)

Effects of predictor variables on relative investment in household children. Ordinal logistic regression model with elastic net penalties.
:::

::: {#sfig-plot-caregiverresponse-coefs}
![](Figures/plot_caregiverresponse_coefs.svg)

Effects of predictor variables on caregiver responses to the last time a child was sad or cried. Ordinal logistic regression model with elastic net penalties.
:::

# Deleted sections

### Child anthropometrics

xxxxxxxxxxDescribe camps brieflyxxxxxxxxxxxxxx - include dates and how we handled those. For cross-sectional analyses, we used the 2023 anthropometric values due to them being closer in time to the caretaker surveys.

All anthropometric measurements were taken twice, in accordance with anthropometric standards [@lohman1988], and then averaged. Before being measured, participants were asked to remove their shoes and heavy articles of clothing. Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that was used in a building with a concrete foundation. All participants were wearing lightweight clothing, so we made no adjustments for clothing weight. Grip strength was measured using a dynamometer (model name), with participants seated/standing? and told to xxxxxxxx. XXXX time was passed between the first and second measurements.

Triceps and subscapular skinfold thicknesses were measured using skinfold calipers. The triceps measurement was taken at the midpoint of the upper arm. The subscapular measurement occurred immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna]. Flexed bicep circumference was measured perpendicular to the long axis of the upper arm. For this measure participants were asked to sit/stand? and position their upper arm perpendicular to their body with their elbow being held at a 90° angle, before flexing their clenched fist towards themselves as much as they could.

We computed each child's height, weight, and BMI Z-scores with respect to the CDC child growth standards [@kuczmarski2002].

For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011]. The resulting variables represent how each child differed from the average child for their given sex and age. We also centered and scaled the skinfold thickness residuals before summing them to create a body fat index.

### Old discussion parts

The effect of age was smallest on child sadness. However, interpretation of this finding is more difficult due to the fact that, unlike crying and temper tantrums, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011]. By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.